Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-07 Manuscript/Manuscript.Rmd
MS edits

**Depolarisation of cPRC is required for activation of downstream interneurons and ciliomotor neurons**

therefore compared calcium imaging of postsynaptic cells (INNOS, INRGW, Ser-h1 and MC) in the cPRC circuit during UV stimulation of cPRC between WT and NOS mutant larvae.

We next quantified the responses of the INNOS and INRGW following a UV/violet stimulus to the cPRCs. To unambiguously identify the cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**. We used the cell-specific neuropeptide markers using anti-RYamide (INNOS) and anti-RGWamide (INRGW) antibodies [@Conzelmann2011] to mark the interneurons. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**. In both wild-type and *NOS*-mutant larvae, INNOS cells showed increase calcium signals during the inhibition phase of the cPRCs **(Figure 5D)**. In addition, INNOS depolarisation was suppressed in NIT-GC2 molphants where cPRC hyperpolarisation was suppressed **(Figure 5E)**. This indicates that cPRC hyperpolarisation is required for INNOS depolarisation. In contrast, calcium signals increased in INRGW cells in wild-type larvae during the second activation phase of cPRCs, a response that was lacking in *NOS*-mutants **(Figure 5F,G)**. This suggests that there is no circuit output to the Ser-h1 and MC cells and the ciliary band in *NOS*-mutants, consistent with the observed behavioural defects **(Figure 5H,I)**.

>>>>

**NO signalling shapes the dynamics of the cPRC circuit**

To investigate how NO and NIT-GC signalling influence the dynamics of the cPRC circuit, we imaged calcium signals from postsynaptic neurons in wild type, mutant and morphant larvae. We were able to image the activity of all neurons in the cPRC circuit (INNOS, INRGW, Ser-h1 and MC). The MC cell was identified based on its position and intrinsic activity. To unambiguously identify all other cells from which we recorded calcium signals, we developed an on-slide immunostaining method **(Figure 5---figure supplement 1)**. We used the cell-specific antibody markers against RYamide (INNOS), RGWamide (INRGW) and serotonin (Ser-h1) [@Conzelmann2011] to immunostain agar-embedded larvae following calcium imaging. Based on the position of the nuclei, we could correlate live and fixed samples at a single-cell precision **(Figure 5A,B)**. Due to the stereotypy of the larvae, we could also identify neurons based on their position and activity in activity-correlation maps **(Figure 5C)**.

We first quantified the responses of the INNOS and INRGW interneurons during 405 nm stimulation of the cPRCs. In both wild type and *NOS*-mutant larvae, INNOS cells showed an increase in calcium during stimulation **(Figure 5D)**. This response was suppressed in NIT-GC2 morphant larvae **(Figure 5E)** revealing an essential role for NIT-GC2-mediated cPRC suppression in INNOS activation. 
In contrast, INRGW cells were initially inhibited, followed by a delayed activation paralleling the second activation phase of cPRCs. This late INRGW response was lacking in *NOS*-mutants **(Figure 5F,G)**. In NIT-GC2 morphants, INRGW cells showed a transient increase in calcium that decayed after light off and a delayed activation was not present.
Next, we imaged calcium signals from Ser-h1 and MC neurons in wild type and *NOS* mutant larvae. Ser-h1 cells showed an activation profile that correlated with cPRC activity, including a reduction in calcium during stimulation followed by rebound, a response that was defective in *NOS* mutants. MC cells showed sustained activation, including a late-phase that was lacking in *NOS* mutants **(Figure 5H,I)**. These data suggest that during 405 stimulation the Ser-h1 cells are inhibited and MC cells are activated, and this regulation is NO-dependent. This pattern is expected to inhibit ciliary activity in the prototroch but not in the other ciliary bands, triggering NO-dependent downward swimming.
